---
title: "20200317_tidy_tuesday"
output: html_document
---

This weeks data is all about the (American) Office's episode ratings. 
One of the variables is of `date` type, I have little experience with this. 
It would be cool to learn more about that. 

```{r}

tuesdata <- tidytuesdayR::tt_load(2020, week = 12)

office_ratings <- tuesdata$office_ratings
schrute_data <- schrute::theoffice

```

## Exploration

### office_ratings

```{r}

skimr::skim(office_ratings)

```

```{r}

library(dplyr)
library(ggplot2)
library(tidyr)
library(scales)

office_ratings %>% 
  dplyr::count(season) %>% 
  ggplot(aes(x=factor(season), y=n)) + 
  geom_bar(stat="identity") +
  ggtitle("Number of episodes per season", 
          subtitle = "Season 4 has far fewer rated episodes")

```

Season 4 has 19 episodes accoring to [Wikipedia](https://en.wikipedia.org/wiki/The_Office_(American_season_4)). 
This season has multiple "double" episodes, meaning that two episode are squished togther, with one name and storyline. 
These are apparently rated as 1 by the IMDB system. 

```{r}

office_ratings %>% 
  ggplot(aes(x=factor(season), y=imdb_rating, fill = factor(season))) + 
  geom_boxplot() + 
  theme(legend.position = "none") + 
  ggtitle("Ratings per season")

office_ratings %>% 
  ggplot(aes(x=factor(season), y=total_votes, fill = factor(season))) + 
  geom_boxplot() + 
  theme(legend.position = "none") + 
  ggtitle("Number of votes per season", 
          subtitle = "Generally, people vote less often for later season.")

office_ratings %>% 
  ggplot(aes(x=total_votes, y=imdb_rating)) + 
  geom_text(aes(label = episode)) + 
  theme(legend.position = "none") + 
  facet_wrap(season ~ . ) +
  ggtitle("#votes vs IMDB score", 
          subtitle = "It appears that more votes means higher score") 

office_ratings %>% 
  group_by(season) %>% 
  mutate(season_n = n()) %>% 
  rowwise() %>% 
  mutate(ep_season =rescale(episode, from=c(1,season_n))) %>%
  ggplot(aes(x=ep_season, y=imdb_rating, group=season, col=factor(season))) +
  geom_line() +
  ggtitle("Progression of score over the season")

office_ratings %>% 
  rowwise() %>% 
  mutate(month = months(air_date)) %>% 
  ungroup() %>% 
  dplyr::count(season, month) %>% 
  spread(key = season, value = n) %>% 
  mutate(month = factor(month, levels = month.name)) %>% 
  arrange(month)

```

### schrute_data

```{r}

skimr::skim(schrute_data)

```

I see that both season and episode are classified as character, let's remedy that. 

```{r}

unique(schrute_data$season)
unique(schrute_data$episode)

schrute_data %>% 
  mutate(season = as.numeric(season)) %>% 
  count(is.na(season))

schrute_data %>% 
  mutate(episode = as.numeric(episode)) %>% 
  count(is.na(episode))

schrute_data <- schrute_data %>%
  mutate(episode = as.numeric(episode), 
         season = as.numeric(season))

```

I have decided to look at the rating of the episode vs the amount of female influence, eg. female director/writer/cast. 
I would also like to look at the number of lines per female character, similar to that of the oscar speeches. 

#### Directors

I'll start with asigning gender to the writers and directors.

```{r}


unique(schrute_data$director) %>% head()

schrute_data %>% 
  select(season, episode, director) %>% 
  unique() %>% 
  filter(grepl(";",director))

schrute_data_directors <- schrute_data %>% 
  select(season, episode, director) %>% 
  unique() %>% 
  filter(!grepl(";",director))

```

For the directors, there are only three episodes with shared director, I am okay with losing those data. 

```{r}

library(stringr)

schrute_data_directors %>% 
  rowwise() %>% 
  mutate(n_spaces = str_count(string = director, pattern = " ")) %>%
  count(n_spaces)

schrute_data_directors %>% 
  rowwise() %>% 
  mutate(n_spaces = str_count(string = director, pattern = " ")) %>%
  filter(n_spaces > 1) 

```

Only Victor Nelli Jr. has a difficult name. 
This will cause the warning that for two rows I provide too few colnames, this is okay. 

```{r}

library(gender)

gender_directors <- schrute_data_directors %>% 
  separate(director, into = c("first_name", "last_name"), sep = " ") %>% 
  pull(first_name) %>% 
  unique() %>% 
  gender()

schrute_data_directors <- schrute_data_directors %>% 
    separate(director, into = c("first_name", "last_name"), sep = " ") %>% 
  left_join(gender_directors[,c(1,4)], by=c("first_name"="name")) 

schrute_data_directors %>% 
  count(gender) %>% 
  ggplot(aes(x=gender, y=n, fill=gender)) + 
  geom_histogram(stat="identity") +
  ggtitle("Number of directors for men and women", 
          subtitle = "I knew it would be bad, but this is extreme")

schrute_data_directors %>% 
    count(season, gender) %>% 
  filter(!is.na(gender)) %>% 
  spread(key=gender, value=n) %>% 
  replace_na(list("male"=0, "female"=0)) %>% 
  mutate(percent_male = male / (male+female), 
         percent_female = 1 - percent_male) %>%
  select(season, percent_male, percent_female) %>%
  gather(key = "gender", value = "percent", -season) %>%
  ggplot(aes(x=factor(season), y=percent, fill=gender)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::percent) +
  ggtitle("Percentage of directors for men and women per season", 
          subtitle = "Arguably it is getting slowly more even..")

```

I feel I need to note here that `NA` means that the name is unconvincingly male or female, not that the director is necessarily non-binary. 

#### Writers

```{r}

unique(schrute_data$writer) %>% head()

schrute_data %>% 
  select(season, episode, writer) %>% 
  unique() %>% 
  rowwise() %>% 
  mutate(n_writers = str_count(string = writer, pattern = ";") + 1) %>% 
  count(n_writers)

schrute_data_writers <- schrute_data %>% 
  select(season, episode, writer) %>% 
  unique() %>% 
  rowwise() %>% 
  mutate(n_writers = str_count(string = writer, pattern = ";") + 1) %>% 
  filter(n_writers == 1) %>% 
  select(-n_writers) 
  
```

I will ignore episodes with multiple writers as well for now. 

```{r}

schrute_data_writers %>% 
  rowwise() %>% 
  mutate(n_spaces = str_count(string = writer, pattern = " ")) %>%
  filter(n_spaces > 1) 

```

Anthony still has a useful first name. 
I will have to ignore his last name. 

```{r}

gender_writers <- schrute_data_writers %>% 
  separate(writer, into = c("first_name", "last_name"), sep = " ") %>% 
  pull(first_name) %>% 
  unique() %>% 
  gender()

schrute_data_writers <- schrute_data_writers %>% 
    separate(writer, into = c("first_name", "last_name"), sep = " ") %>% 
  left_join(gender_writers[,c(1,4)], by=c("first_name"="name")) 

```

```{r}

schrute_data_writers %>% 
  count(gender) %>% 
  ggplot(aes(x=gender, y=n, fill=gender)) + 
  geom_histogram(stat="identity") +
  ggtitle("Number of directors for men and women", 
          subtitle = "This is already far better than the directors")

schrute_data_writers %>% 
    count(season, gender) %>% 
  filter(!is.na(gender)) %>% 
  spread(key=gender, value=n) %>% 
  replace_na(list("male"=0, "female"=0)) %>% 
  mutate(percent_male = male / (male+female), 
         percent_female = 1 - percent_male) %>%
  select(season, percent_male, percent_female) %>%
  gather(key = "gender", value = "percent", -season) %>%
  ggplot(aes(x=factor(season), y=percent, fill=gender)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::percent) +
  ggtitle("Percentage of directors for men and women per season", 
          subtitle = "I suppose this is sortof stable, which is a little dissapointing.")

```

#### Female characters

```{r}

character_genders <- schrute_data %>% 
  separate(character, c("first_name", "second_name"), sep = " ") %>% 
  pull(first_name) %>% 
  unique() %>% 
  gender()

```

There are a ton of characters that do not have a neat first_name last_name structure. 
That is okay. 

```{r}

schrute_data %>% 
  select(character) %>% 
  unique() %>% 
  separate(character, c("first_name", "second_name"), sep = " ") %>% 
  left_join(character_genders[,c(1,4)], by=c("first_name"="name")) %>% 
  filter(is.na(gender)) %>% 
  count(first_name) %>% 
  arrange(desc(n))

characters_interest <- schrute_data %>% 
    select(season, character) %>% 
  unique() %>% 
  count(character) %>% filter(n > 2) %>%
  left_join(character_genders[,c(1,4)], 
            by=c("character"="name")) %>% 
  filter(!is.na(gender)) %>%
  pull(character)

schrute_data %>% 
  select(season, episode, character, text) %>% 
  filter(character %in% characters_interest) %>%
  left_join(character_genders[,c(1,4)], by=c("character"="name")) %>%
  group_by(season, episode) %>% 
  count(gender) %>% 
  spread(key = gender, value = n) %>% 
  mutate(percent_male_lines = male / (male + female),
         percent_female_lines = 1 - percent_male_lines) %>% 
  select(season, episode, percent_male_lines, percent_female_lines) %>% 
  gather(key = gender, value = percent, -season, -episode) %>%
  ggplot(aes(x=factor(season), y=percent, col=gender)) +
  geom_jitter()

```

There are some interesting things in here, but let's focus on named characters that recur in at least three seasons. 
I selected characters who are known by only their first name and appear in at least three seasons. 
To me this indicates they are important enough to have sufficient lines to be interesting to compare. 

```{r}

character_genders %>% 
  filter(name %in% characters_interest) %>% 
  count(gender) %>% 
  ggplot(aes(x = gender, y=n, fill=gender)) + 
  geom_histogram(stat = "identity") +
  ggtitle("Number of men and women in my selection", 
          subtitle = "Why is this not surprising anymore")


```

```{r}

schrute_data %>% 
  filter(character %in% characters_interest) %>% 
  rowwise() %>% 
  mutate(n_words_line = str_count(string = text, pattern = " ") + 1) %>%
  select(season, episode, writer, character, n_words_line) %>%
  left_join(character_genders[,c(1,4)], by=c("character"="name")) %>%
  ggplot(aes(x=gender, y=log2(n_words_line))) + 
  geom_violin() +
  ggtitle("Number or words per line for each gender" + 
            subtitle = "When speaking, both genderes get a similar number of words in a line")


```


### Plain correlations

```{r}

office_ratings %>% 
  select(season, episode, imdb_rating) %>%
  left_join(schrute_data_writers, by=c("season", "episode")) %>%
  na.omit() %>% 
  ggplot(aes(x=gender, y=imdb_rating)) + 
  ggbeeswarm::geom_beeswarm() +
  ggtitle("Ratings for episodes by female vs male writers")

```


